---
title: "p8105_hw2_ts3349"
author: "Tessa Senders"
date: "9/25/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(readxl)
library(readr)
```


## Problem 1

Read the Mr. Trashwheel data set.

```{r trashwheel}
trashwheel_df <- read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                           sheet = "Mr. Trash Wheel", 
                           range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls), 
         sports_balls = as.integer(sports_balls))
  
head(trashwheel_df)
```



Read the precipitation data sets for 2017 and 2018.

```{r precip}
precip_2018 <- read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                           sheet="2018 Precipitation", 
                           skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 <- read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                           sheet="2018 Precipitation", 
                           skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>%
  relocate(year)

head(precip_2017)
```



```{r bind}
month_df <- 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df <- bind_rows(precip_2018,precip_2017) 

precip_df <- left_join(precip_df, month_df, by = "month") %>% 
  select(!month) %>% 
  mutate(month_name = tolower(month_name))

head(precip_df)
```
This data set contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland.  As trash enters the inner harbor, the trashwheel collects that trash and stores it in a  dumpster.  The dataset contains information on year, month, and trash collected including like plastic bottles, chip bags, and sports balls.  There are a total of `r nrow(trashwheel_df)` rows in the final data set.  The median number of sports balls for example is `r median(pull(trashwheel_df, sports_balls))`.  There are additional data sheets which include month precipitation data.


## Problem 2

Read the NYC Transit data.

```{r nyc transit data read}
nyc_transit_df <- read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>%
  mutate(entry = recode(entry, YES = TRUE, NO = FALSE))

  
head(nyc_transit_df)
#View(nyc_transit_df)
```

This data set contains information about the New York City transit system (the Metropolitan Transportation Authority of the State of New York), specifically the NYC subway stations and their entrances and exits.  The data set contains information on station names, station latitude/longitude, lines, routes, ability to enter, entrance type, and whether the station is ADA accessible.  In cleaning the data, many columns were removed including division, column regarding staffing, columns with notes, corner location, and columns regarding entrance location.  Only the variables previously mentioned were selected and all variable names were cleaned to follow snake case.  The entry variable was converted from a character variable to a logical variable where "YES" was converted to "TRUE" meaning if the station has an entry the entry variable is "TRUE" and if the station does not have an entry it is "FALSE" instead of "NO".  The data set has `r nrow(nyc_transit_df)` rows and `r ncol(nyc_transit_df)` columns.  The data is not tidy yet because each different route is a column.  Route is a variable and should have one column containing the different routes.


Answer Questions about the NYC Transit data:
```{r}
nyc_stations_df <- nyc_transit_df %>% 
  unite(distinct_stations, c("line", "station_name"), sep = "_") %>% 
  group_by(distinct_stations) %>%
  count()

head(nyc_stations_df)
nrow(nyc_stations_df)

#465

nyc_stations_ADA <- nyc_transit_df %>% 
  unite(distinct_stations, c("line", "station_name"), sep = "_") %>% 
  group_by(distinct_stations, ada) %>%
  summarize()

head(nyc_stations_ADA)
sum(pull(nyc_stations_ADA, ada))
#84

nyc_stations_entrance <- nyc_transit_df %>% 
  unite(distinct_stations, c("line", "station_name"), sep = "_") %>%
  filter(vending == "NO") %>%
  group_by(distinct_stations, entry) %>%
  summarize()

head(nyc_stations_entrance)
sum(pull(nyc_stations_entrance, entry)) / nrow(nyc_stations_entrance)
#0.3873874

  

  
```

There are `r nrow(nyc_stations_df)` distinct stations.  Of all the stations, `r sum(pull(nyc_stations_ADA, ada))` are ADA compliant.The proportion of stations entrance/exits without vending that allow entry is `r sum(pull(nyc_stations_entrance, entry)) / nrow(nyc_stations_entrance)`.
   













