---
title: "p8105_hw2_ts3349"
author: "Tessa Senders"
date: "9/25/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(readxl)
```


## Problem 1

Read the Mr. Trashwheel data set.

```{r}
trashwheel_df <- read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                           sheet = "Mr. Trash Wheel", 
                           range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls), 
         sports_balls = as.integer(sports_balls))
  
head(trashwheel_df)
```



Read the precipitation data sets for 2017 and 2018.

```{r precip}
precip_2018 <- read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                           sheet="2018 Precipitation", 
                           skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 <- read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                           sheet="2018 Precipitation", 
                           skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>%
  relocate(year)

head(precip_2017)
```



```{r bind}
month_df <- 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df <- bind_rows(precip_2018,precip_2017) 

precip_df <- left_join(precip_df, month_df, by = "month") %>% 
  select(!month) %>% 
  mutate(month_name = tolower(month_name))

head(precip_df)
```
This data set contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland.  As trash enters the inner harbor, the trashwheel collects that trash and stores it in a  dumpster.  The dataset contains information on year, month, and trash collected including like plastic bottles, chip bags, and sports balls.  There are a total of `r nrow(trashwheel_df)` rows in the final data set.  The median number of sports balls for example is `r median(pull(trashwheel_df, sports_balls))`.  There are additional data sheets which include month precipitation data.




