---
title: "p8105_hw2_ts3349"
author: "Tessa Senders"
date: "9/25/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(readxl)
library(readr)
library(lubridate)
```


## Problem 1

Read the Mr. Trashwheel data set.

```{r trashwheel}
trashwheel_df <- read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                           sheet = "Mr. Trash Wheel", 
                           range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls), 
         sports_balls = as.integer(sports_balls))
  
head(trashwheel_df)
```



Read the precipitation data sets for 2017 and 2018.

```{r precip}
precip_2018 <- read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                           sheet="2018 Precipitation", 
                           skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 <- read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                           sheet="2018 Precipitation", 
                           skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>%
  relocate(year)

head(precip_2017)
```



```{r bind}
month_df <- 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df <- bind_rows(precip_2018,precip_2017) 

precip_df <- left_join(precip_df, month_df, by = "month") %>% 
  select(!month) %>% 
  mutate(month_name = str_to_lower(month_name))

head(precip_df)
```
This data set contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland.  As trash enters the inner harbor, the trashwheel collects that trash and stores it in a  dumpster.  The dataset contains information on year, month, and trash collected including like plastic bottles, chip bags, and sports balls.  There are a total of `r nrow(trashwheel_df)` rows in the final data set.  The median number of sports balls for example is `r median(pull(trashwheel_df, sports_balls))`.  There are additional data sheets which include month precipitation data.


## Problem 2

Read the NYC Transit data.

```{r nyc transit data read}
nyc_transit_df <- read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>%
  mutate(entry = recode(entry, YES = TRUE, NO = FALSE))

  
head(nyc_transit_df)
#View(nyc_transit_df)
```

This data set contains information about the New York City transit system (the Metropolitan Transportation Authority of the State of New York), specifically the NYC subway stations and their entrances and exits.  The data set contains information on station names, station latitude/longitude, lines, routes, ability to enter, entrance type, and whether the station is ADA accessible.  In cleaning the data, many columns were removed including division, column regarding staffing, columns with notes, corner location, and columns regarding entrance location.  Only the variables previously mentioned were selected and all variable names were cleaned to follow snake case.  The entry variable was converted from a character variable to a logical variable where "YES" was converted to "TRUE" meaning if the station has an entry the entry variable is "TRUE" and if the station does not have an entry it is "FALSE" instead of "NO".  The data set has `r nrow(nyc_transit_df)` rows and `r ncol(nyc_transit_df)` columns.  The data is not tidy yet because each different route is a column.  Route is a variable and should have one column containing the different routes.


Answer Questions about the NYC Transit data:
```{r}
nyc_stations_df <- nyc_transit_df %>% 
  unite(distinct_stations, c("line", "station_name"), sep = "_") %>% 
  group_by(distinct_stations) %>%
  count()

head(nyc_stations_df)
nrow(nyc_stations_df)

#465

nyc_stations_ADA <- nyc_transit_df %>% 
  unite(distinct_stations, c("line", "station_name"), sep = "_") %>% 
  group_by(distinct_stations, ada) %>%
  summarize()

head(nyc_stations_ADA)
sum(pull(nyc_stations_ADA, ada))
#84

nyc_stations_entrance <- nyc_transit_df %>% 
  unite(distinct_stations, c("line", "station_name"), sep = "_") %>%
  filter(vending == "NO") %>%
  group_by(distinct_stations, entry) %>%
  summarize()

head(nyc_stations_entrance)
sum(pull(nyc_stations_entrance, entry)) / nrow(nyc_stations_entrance)
#0.3873874

  

  
```

There are `r nrow(nyc_stations_df)` distinct stations.  Of all the stations, `r sum(pull(nyc_stations_ADA, ada))` are ADA compliant.The proportion of stations entrance/exits without vending that allow entry is `r sum(pull(nyc_stations_entrance, entry)) / nrow(nyc_stations_entrance)`.
   

Answer questions regarding the A train:
```{r reformat NYC transit data}
nyc_transit_df_tidy <- nyc_transit_df %>%
  mutate(route8 = as.character(route8)) %>%
  mutate(route9 = as.character(route9)) %>%
  mutate(route10 = as.character(route10)) %>%
  mutate(route11 = as.character(route11)) %>%
  pivot_longer(
    route1:route11,
    names_to = "route_name", 
    values_to = "route_number")

nyc_transit_A_train_df <- nyc_transit_df_tidy %>% unite(distinct_stations, c("line", "station_name"), sep = "_") %>% 
  filter(route_number == "A") %>%
  group_by(distinct_stations, route_number, ada) %>%
  summarize()
  
nrow(nyc_transit_A_train_df)
sum(pull(nyc_transit_A_train_df, ada))

#View(nyc_transit_df_tidy)
#View(nyc_transit_A_train_df)
```

`r nrow(nyc_transit_A_train_df)` distinct stations serve the A train.  Of the stations that serve the A train, `r sum(pull(nyc_transit_A_train_df, ada))` are ADA compliant.

## Problem 3

Read in the FiveThirtyEight pols_month data.
```{r read fivethirtyeight pols_month data}
pols_month_df <- read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>%
  mutate(month_2 = month(ymd(mon), label = TRUE, abbr = TRUE)) %>%
  separate(mon, into=c("year", "month", "day"), sep = "-") %>%
  mutate(month = str_to_lower(month_2)) %>%
  select(!month_2) %>%
  mutate(president = case_when(prez_gop == 1 ~ "gop", prez_gop == 2 ~ "gop_ford_after_nixon", prez_gop == 0 ~ "dem")) %>%
  select(!prez_gop & !prez_dem & !day) 
```



pols-month.csv, unemployment.csv, and snp.csv. Our goal is to merge these into a single data frame using year and month as keys across datasets.

clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns

WHICH IS MONTH AND WHICH IS DAY PLEAE FIX!
```{r fivethirtyeight snp data}
snp_df <- read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(month_2 = month(mdy(date), label = TRUE, abbr = TRUE)) %>%
  separate(date, into=c("month", "day", "year"), sep = "/") %>%
  mutate(month = str_to_lower(month_2)) %>%
  select(!month_2 & !day)  %>%
  relocate(year)

#View(snp_df)
```
Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.
```{r fivethirtyeight unemployment data}
unemployment_df <- read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month", 
    values_to = "unemployment_rate") %>%
  mutate(year = as.character(year))

#View(unemployment_df)
```
Join the datasets by merging snp into pols, and merging unemployment into the result:
```{r join fivethirtyeight data}
pols_snp_df <- left_join(pols_month_df, snp_df, by=c("year" = "year", "month" = "month"))
View(pols_snp_df)

pols_snp_unemployment_df <- left_join(pols_snp_df, unemployment_df, by=c("year" = "year", "month" = "month"))
View(pols_snp_unemployment_df)
```














